---
title: "Programming for Analytics in R- Nolan's Book Club"
output:
  html_document: default
  pdf_document: default
  date: "Fall 2023"
---
<!-- name of chunk = set_global_options and include chunk in doc -->
```{r set_global_options, echo=TRUE}
 knitr::opts_chunk$set(echo = TRUE)
```
The raw data file for this project is full of problems. There are missing values, outliers and invalid values (e.g., negative revenue). There are extra commas at the end of each line that cause the creation of empty nonsense variables and a few lines are missing a comma to separate the last two variables. Many lines also have zero purchases for all products.

Before doing any analytics, we must first scrub the raw data to solve these problems. We make use of the **dplyr** and **tidyr** packages for many of the scrubbing tasks.

```{r echo=TRUE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)


cbc <- as.data.frame(read_excel("/Users/nolancutler/Desktop/Programming for Analytics/Charles_BookClub_modified.xls"))
###Importing data from Excel file
cbc <- cbc[,c(1:18)]
str(cbc)
summary(cbc) 
### List variables, show type
```
#### ***Count the unique values of each variable.***
```{r}
cbc_counts <- cbc %>% summarise_all(n_distinct)
cbc_counts
```

#### ***Convert Gender and Florence to Factors***
```{r}
cbc$Gender <- factor(cbc$Gender, levels = c(0,1), labels = c("Female", "Male"))
cbc$Florence <- factor(cbc$Florence, levels = c(0,1), labels = c("No", "Yes"))
```
#### Convert M and R variables to number values
```{r}
cbc$M = as.numeric(cbc$M, na.rm = TRUE)
cbc$R = as.numeric(cbc$R, na.rm = TRUE)
```
#### Identify the numeric columns for which outlier detection is desired
```{r}
outvars <- c("M", "R", "F", "FirstPurch")
```


#### ***Find outliers and set them to missing (NaN)***
Note the use of the *anonymous* function in the following code:
```{r}

cbc[outvars] <- data.frame(lapply(cbc[outvars], function(x) {
  ifelse((x < 0) | x > (mean(x, na.rm = TRUE) + 3*sd(x, na.rm = TRUE)), NA, x) 
  }))

```

#### ***Summary also counts the number of missing values***
```{r}

summary(cbc)

```
####Identify variables for which imputation of missing values is desired
```{r}
missvars <- c(outvars)
```

#### ***Impute missing values of columns with missing values***
Here's another *anonymous* function use:
```{r}
cbc[missvars] <- data.frame(lapply(cbc[missvars], function(x) {
  ifelse(is.na(x), mean(x, na.rm = TRUE), x) 
  }))

summary(cbc)
```

#### ***Delete rows for which there are no books purchased.***
```{r}
book_rents <- (cbc[8:17])
cbc_no_zeroes <- cbc %>% filter 
nrow(cbc)
nrow(cbc_no_zeroes)
```

#### Sum the purchases of each book type.
```{r}
cbc_sums <- cbc %>% summarise(across(c(ChildBks:ItalArt), sum))
cbc_sums
```

### ***Histogram plot of numeric variables***

```{r}
library(psych)
multi.hist(cbc[, 8:17],nrow=2,ncol=2, global = FALSE)
```

### ***Bar plot of book type sums***

```{r}
cbc_pivot <- pivot_longer(cbc_sums, cols=c(ChildBks:ItalArt))
names(cbc_pivot) <- c("Type", "Sum")

ggplot(data = cbc_pivot, aes(y = Sum, x = Type)) +
  geom_bar(stat = 'identity')
```


### ***A custom function for calculating 4 moments (mean, stdev, kurt)***

```{r}
library(e1071)
#browser()
calcfourstats <- function(x) {
  mu <- round(mean(x), 2)
  sigma <- round(sd(x), 2)
  skew <- round(skewness(x), 3)
  kurt <- round(kurtosis(x), 2)
  result <- data.frame(mu, sigma, skew, kurt)
  return(result)
}

results <- calcfourstats(cbc_no_zeroes$M)
results <- rbind(results, calcfourstats(cbc_no_zeroes$R))
#browser()
results <- rbind(results, calcfourstats(cbc_no_zeroes$F))
results <- rbind(results, calcfourstats(cbc_no_zeroes$FirstPurch))
varList <- names(cbc_no_zeroes[outvars])
print(varList)
rownames(results) <- varList
print(results)
```

### Creating RFM factors
#### Calculate HML cutoffs for RFM
```{r}
cbc_rfm <- data.frame(lapply(cbc_no_zeroes[c("R", "F", "M")], 
  function(x) {
    quantile(x, probs = c(0.33, 0.66, 1.0), na.rm = TRUE) 
    }))
```

Verify results and test subsetting    
```{r}
cbc_rfm
cbc_rfm["33%", "M"] #What is the 33rd percentile of M?
```

### Create three new variables for HML quantiles of RFM variables
```{r}
library(dplyr)
cbcRFM <- cbc_no_zeroes %>%
  #Grouped all inside mutate
  mutate(rRFM = if_else(R <= cbc_rfm["33%", "R"], "L",
                        if_else(R >= cbc_rfm["66%", "R"], "H", "M"))) %>%
  mutate(fRFM = if_else(F <= cbc_rfm["33%", "F"], "L",
                        if_else(F >= cbc_rfm["66%", "F"], "H", "M"))) %>%
  mutate(mRFM = if_else(M <= cbc_rfm["33%", "M"], "L",
                        if_else(M >= cbc_rfm["66%", "M"], "H", "M")))
```
Convert the new HML variables into ordered factors
```{r}
cbcRFM[c("rRFM", "fRFM", "mRFM")] <- data.frame(lapply(cbcRFM[c("rRFM", "fRFM", "mRFM")], 
  function(x) {
    factor(x, c("L", "M", "H"), ordered = TRUE)
  }))

head(cbcRFM)
str(cbcRFM)

sumTable <- cbcRFM %>% 
  group_by(rRFM, fRFM, mRFM) %>%
  summarize(meanM = round(mean(M), 2))

sumTable
```

#### Make three tables, one for each level of factor mRFM

```{r, echo=TRUE, message=FALSE, warning=FALSE}
for (i in c("L", "M", "H")) {
  shortTable <- xtabs(meanM ~ rRFM + fRFM, sumTable %>% filter(mRFM == i)) 
    print(paste('Monetary Value Segment =', i))
    print(shortTable)
    cat("\n") # Add a blank line between tables
    
} 
  
```

### ***Median monetary value per visit by gender***

```{r}
visitValue <- cbcRFM %>% 
  group_by(factor(Gender, labels = c("Female", "Male"))) %>%
  summarize(medianM = round(median(M / F), 2))

visitValue
```

#### ***Bubble plot of M by R by Gender***

```{r}
ggplot(cbcRFM, aes(x = R, y = M, col = factor(Gender, labels = c("Female", "Male")), size = FirstPurch)) +
  geom_point(alpha = .20) +
  
  labs(x = "Recency", y = "Monetary Value") +
  facet_wrap(~ factor(Gender, labels = c("Female", "Male")), labeller = label_parsed) +
  theme(legend.position = "bottom", legend.box = "vertical", 
        legend.key = element_rect(colour = 'white', fill = 'white'))
```

