---
title: "Programming for Analytics- Tripadvisor"
output:
  html_document: default
  pdf_document: default
  date: "Fall 2023"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) ### Importing tidyverse instead of individual packages
```
### Data import
```{r}
hotels <- read_csv("/Users/nolancutler/Desktop/Programming for Analytics/Assignment 2/hotels.txt",
    col_types = cols(hotel_code = col_character(),
                     location_code = col_character()))
   ### Hotel codes and locations should be char values- does not make sense to perform mathematical operations
```
#### Describing imported data
```{r}
head(hotels)
str(hotels)
summary(hotels)
```
From the summary we can see that there are outliers and many missing values in a number of numeric variables.

### Business Questions

#### How many observations are in the dataset? Observations are rows, therefore nrow() will get the job done.

```{r}
nrow(hotels)
```

#### How many stays per each property type?

```{r}
hotels %>%
  group_by(hotel_type) %>%
  tally()
```

#### Outliers
Identify columns for which we wish to find outliers and set to NA.

```{r}
outvars <- c("rooms", "bubble_one", "bubble_two", "bubble_three", "bubble_four", "bubble_five",
             "page_position", "out_of", "reviews", "reviews_per_room", "traffic", "traffic_per_room",
             "OTA_region_rate")

hotels[outvars] <- data.frame(lapply(hotels[outvars], function(x) {
  ifelse(x %in% boxplot.stats(x)$out, NA, x) }))

summary(hotels)

```

#### Missing values
```{r}

missvars <- c("rooms", "bubble_rating", "bubble_one", "bubble_two", "bubble_three", "bubble_four",
              "bubble_five", "page_position", "out_of", "reviews", "domestic_reviews", "international_reviews",
              "reviews_per_room", "management_response_rate", "traffic", "traffic_per_room", "OTA_region_rate")

hotels[missvars] <- data.frame(lapply(hotels[missvars], function(x) {
  ifelse(is.na(x), mean(x, na.rm = TRUE), x) }))

summary(hotels)

```

#### Using ggplot2 to prooduce a boxplot chart of traffic by hotel_type.

```{r}
ggplot(hotels, aes(x = hotel_type, y=traffic))+
  geom_boxplot()+
  facet_grid(.~hotel_type)+
  labs(title = "Traffic by Property Type", x ="property type")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

Note that traffic is highly skewed. See what happens if we transform traffic with the log function:

```{r}
hotels %>% 
  mutate(log_traffic = log(traffic)) %>%
  ggplot(aes(x = hotel_type, y = log_traffic)) +
    geom_boxplot()+
    facet_grid(.~hotel_type) +
    labs(title = "Traffic by Property Type", x ="Property Type")+
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5))
```

#### Finding number of business subscriptions

```{r}
hotels %>%
  filter(subscriber == 1) %>%
  group_by(subscriber) %>%
  tally()
```
There are 46935 Business subscribers.

#### Calculating proportion of subscribers for each country and identifying top five countries by subscriber proportion.

```{r}
hotels %>%
  filter(hotel_type == "Hotel") %>%
  group_by(country_name) %>%
  summarize(subscriber_prop = mean(subscriber, na.rm = TRUE)) %>%
  arrange(desc(subscriber_prop)) %>%
  head(5)
```

#### Ocean City has the most readers of reviews from China.

```{r}
# Find the country code for China:
hotels %>%
  filter(country_name == "China")


hotels %>%
  filter(country_name == "United States" & 
           (visitor_country1 == "CHN" | visitor_country2 == "CHN" | visitor_country3 == "CHN")) %>%
  count(location_name, sort = TRUE)
```

#### What proportions of those Chinese review readers of Ocean City prefer Hotel, B&B and "other" for their stays?

```{r}
hotels %>%
  filter(location_name == "Ocean City" & 
         (visitor_country1 == "CHN" | visitor_country2 == "CHN" | visitor_country3 == "CHN")) %>%
  group_by(hotel_type) %>%
  summarise(n = n()) %>% 
  ungroup() %>%
  complete(hotel_type = c("Hotel", "B & B", "Other"), fill = list(n = 0)) %>%
#### Using Complete to show reservations of "other", even though they are at 0.
  mutate(freq = round(n / sum(n), 2))
```